Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by XXXXXXX research grants.}

### 2015-10-22 latex/article.tex
reverting destructive commit >_<

\begin{figure}[t]{\small
    \begin{center}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/integral2D_2}
          \put(42,28){$\times$}
          \put(46,31){$\vx$}
          \put(55,45){$\Ball{R}{\vx}$}
          \put(1,10){$\dS$}
          \put(20,10){$\Shape$}
      \end{overpic}}~~~~~~
      \subfigure[]{\begin{overpic}[width=5cm]{figs/notation_2}
          \put(44.5,25){$h \cdot \vz$}
          \put(55,45){$\Ball{R}{\vx}$}
          \put(42,28.5){$\times$}
          \put(45,32.5){$\vx$}
          \put(45,16.5){$h$}
          %\put(88,39){$\Proj(p)$}
          \put(3,11){$\partial \Shape$}
          \put(15,7){$\partial \Body{\DigShape}{h}$}
      \end{overpic}}
      \vspace{-0.4cm}
    \end{center}}
    \caption{Integral invariant computation $(a)$ and
      notations $(b)$ in dimension 2 \cite{CVIU2014}.\label{fig:notations}}
\end{figure}

@@

\begin{figure}[!htbp]
  \begin{center}
    \subfigure{}{\includegraphics[width=0.9\textwidth]{figs/morton}}
    %%\subfigure{}{\includegraphics[height=0.17\textheight]{figs/partitioning}}
    %%\subfigure{}{\includegraphics[height=0.17\textheight]{figs/quadtree_activefront}}
  \end{center}
\caption{Morton codes associated to cells of a linear quatree. Each
   morton code of a child cell is obtained by adding a suffix to its
  parent code \emph{(left)}. The adaptive representation consists of
  quadtree cells whose depth is view point dependent
  \emph{(middle)}. Finally, adaptive Marching-Cubes is used to
  generate the triangulation \emph{(right)}.}
\label{fig_quadtree_partitionning}
\end{figure}

@@

As illustrated in Figure
\ref{fig_quadtree_partitionning}-\emph{middle}, we propose a viewpoint
dependent criterion to decide if a cell needs to be refined: the
closer we are to the camera, the finer the cells are. In addition to
the distance criterion, we define an angular criterion to prevent refining cells outside the camera frustrum. Using a criterion based on the distance between a cell and the camera is well suited to GPU since it can be evaluated independently on all cells. Hence, we can efficiently determine if a cell must be subdivided or not. Figure \ref{fig_lod_octree}
illustrates the level of details (LoD for short) criterion we
use. More precisely, in dimension 2, if $\alpha$ denotes the viewing
angle, an object at distance $d$ from the camera with diameter greater than
 $2\cdot d\cdot\tan(\alpha)$
 is projected into more than one pixel
(see Figure~\ref{fig_lod_octree}-\emph{left}). Hence, the distance
criterion is based on the ratio (\emph{visibility ratio} in the
following)between the cell diameter $l(c)$ (power of 2 depending on
the depth), and
$2\cdot d_c\cdot\tan(\alpha)$, $d_c$ being the distance between the cell center and the camera. For a given cell $c$,
split and merge decision are based on this visibility ratio:
\begin{itemize}
\item $c$ is split if its children cells have a visibility ratio
  greater than  constant $k$\,;
\item $c$ and its sibling cells are merged if their parent cell $c'$
  have av visibility ratio lower than $k$\;
\item otherwise, the cell $c$ stays for the next frame.
\end{itemize}
%
\begin{figure}[!h]
  \begin{center}
    \begin{overpic}[width=3.5cm]{figs/criterion}
      \put(33,40){$\alpha$}
      \put(50,45){$d_c$}
      \put(95,65){$l(c)$}
    \end{overpic}~~~~~
    \includegraphics[width=3.5cm]{viewlod2_small}~~~~~
    \includegraphics[width=3cm]{subdivision}
  \end{center}
  \caption{Notations \emph{(left)} and adaptive meshing in dimension 2 using the LoD distance and
    angular criterion \emph{(middle)}. Based on LoD criterion
    evaluation, the cell buffer is
  updated between each frame \emph{(right)}.}
  \label{fig_lod_octree}
\end{figure}
%
In addition to the distance criterion, we add an angular term in order
to prevent subdivision of cells far from the camera frustrum (see
Figure~\ref{fig_lod_octree}-\emph{middle}). We do not go into the
details but please note that stay, split and merge decisions of all cells are computed in
parallel from all cell morton codes.

@@

To implement the integration on $\Ball{R}{\vx} \cap \Shape$
(Fig.~\ref{fig:approx}-$a$), a naive approach consists in scanning all
digital points inside such domain (Fig.~\ref{fig:approx}-$b$) and then
to estimate the geometrical moment as the sum of the geometrical
moments of each elementary cubes lying in the intersection. On GPU,
\emph{mipmap} textures are used to obtain a multi-resolution information: if the
input binary object is stored in a 3D texture, GPU hardware can
construct a multi-resolution pyramid such that 8 neighboring voxel
intensities at level $l$ are averaged to define the voxel value at
level $l+1$. If the level 0 corresponds to the binary input object, at
a given level $l$, a voxel $(x,y,z)$ contains the fraction of $\DSh$
belonging to the cube of center $(x,y,z)$ and edge length $2^l$. As a
consequence, we can approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by
considering mipmap values at a given resolution $l$
(Fig.~\ref{fig:approx}-$c$). In this case, errors only occurs for
cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$.
\begin{figure}
  \begin{center}
    \subfigure[]{\includegraphics[width=4cm]{figs/approx1}}
    \subfigure[]{\includegraphics[width=4cm]{figs/approx-reg-1}}
    \subfigure[]{\includegraphics[width=4cm]{figs/approx-reg-2}}
  \end{center}
  \caption{Integral computation with an area estimation from pixel
    counting $(b)$ or from densities (represented by different
    grayvalues) in a mipmap approximation of $\Shape$.}
  \label{fig:approx}
\end{figure}

Using the hierachical nature of the mipmap texture, we could also
consider hierarchical decompositions of either $\Ball{R}{\vx}$
(Fig.~\ref{fig:approx2}-$a$) or $\Ball{R}{\vx}\cap\Shape$
(Fig.~\ref{fig:approx2}-$b$). The idea would be to decompose the
regions into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities. Note that the scenario in
Fig.~\ref{fig:approx2}-$b$, is clearly the best approach: the
volume is exact  (\wrt the regular grid case) since geometrical moments
are additive quantities, and the number of texture fetches is  optimal
at this point.

\begin{figure}
  \begin{center}
    \subfigure[]{\includegraphics[width=4cm]{figs/approx-bh}}
    \subfigure[]{\includegraphics[width=4cm]{figs/approx-ada}}
  \end{center}
  \caption{Adaptive approximation of the integration using a
    hierarchical decomposition of the ball $\Ball{R}{\vx}$ $(a)$, or a decomposition of $\Ball{R}{\vx}\cap\Shape$.}
  \label{fig:approx2}
\end{figure}

@@

\paragraph{Hierarchical Decomposition.}  As discussed above, best
results in terms of texture fetch can be obtained by the approach
presented in Fig.~\ref{fig:approx2}-$b$. At each vertex, the
computation algorithm would consist in a recursive decomposition of
the domain such that a cell is subdivided if it does not belong to
$\Ball{R}{\vx}\cap\Shape$. However, on GPU hardware, this approach is
highly inefficient since the hardware optimizes the parallelism only
when the micro-program described in the shader has a predictive
execution flow. Hence, implementing the recursive algorithm (or its
de-recursified version) requires a lot of branches (conditional
\emph{if} tests) in the flow. Instead of implementing this approach,
we have considered a fast hierarchical approximation following
Fig.~\ref{fig:approx2}-$a$: for a given radius $R$, we precompute a
hierachical decomposition of $B_R$. Such decomposition is made of
octree cells at different resolutions. At a given MC vertex $\vx$, the
algorithm becomes simple since we just scan the ball cells and
estimate the area (or other moments) from the mipmap values associated
to this cell. Note that cells in the $B_R$ decomposition may not be
aligned wuth mipmap cells. However, we can use GPU mipmap
functionalities which interpolates mipmap values if we probe at
non-discrete positions. For a given radius $R$, the expected number of
cells on $B_R$ is in $O(\log{R})$. Implementation details on the
hierarchical octree representation can be found in the supplementary
material (see Section \ref{sec:discussion}).

@@

We first evaluate the curvature estimation on a full-resolution
geometry obtained by disabling the LoD criterion at the mesh
generation step.

\begin{table}
  \begin{center}
    \begin{tabular}{@{}lrrrrrrrrrrrr@{}}
      \toprule
      & \multicolumn{6}{c}{$R=$}   & \multicolumn{6}{c}{$R=$}\\
      \cmidrule(r){2-12}
    Quantities & ~~~H & $l_4$ & $l_3$ & $l_2$ & $l_1$ & $l_0$& ~~~H & $l_4$ & $l_3$ & $l_2$ & $l_1$ & $l_0$\\
      \midrule
      Number of texture fetches & & & & & & & & & & & &\\
      Time (in ms) & & & & & && & & & & &\\
      $L_\infty$ error (\wrt $l_0$) & & & & & && & & & & &\\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{object, size, radius, full res, num of vertices (timings
    only consider the curvature estimation step).\label{tab:full-res-stat}}
\end{table}

\begin{figure}
  \begin{center}
    \subfigure[]{\framebox{octa $l_4$}}
   \subfigure[]{\framebox{octa $l_3$}}
   \subfigure[]{\framebox{octa $l_2$}}
   \subfigure[]{\framebox{octa $l_1$}}
   \subfigure[]{\framebox{octa $l_0$}}
  \end{center}
  \caption{fig multiresolution, courbure moyenne par ex}
  \label{fig:refine}
\end{figure}
\todo[inline]{blabla experiments}

@@

\begin{figure}
  \begin{center}
    \subfigure[]{\framebox{K}}
   \subfigure[]{\framebox{H}}
   \subfigure[]{\framebox{DIR1}}
   \subfigure[]{\framebox{DIR2}}
   \subfigure[]{\framebox{normales}}\\
    \subfigure[]{\framebox{K}}
   \subfigure[]{\framebox{H}}
   \subfigure[]{\framebox{DIR1}}
   \subfigure[]{\framebox{DIR2}}
   \subfigure[]{\framebox{normales}}

  \end{center}
  \caption{fig illustration complete (adaptative, ...) plusieurs
    objets.. + dynamique.}
  \label{fig:full}
\end{figure}

\todo[inline]{blabla experiments}

For dynamic datasets, we consider a time varying impicit
surface. Since it would have been to expensive to update the a mipmap texture
of densities at each time step, we simply evaluate the implicit
expression at each vertex of a cell to decide if such cell is included
or not into $\Ball{R}{\vx}\cap \Shape$.
\todo[inline]{blabla experiments}

@@

In this article, we have propose a full data parallel framework on GPU
hardware which combines an adaptive isosurface construction from
digital data with a curvature tensor estimation at each vertex. Using
this approach, we can explore in realtime different curvature
measurements (mean, Gaussian, principal directions) with different
ball radii on potentially large dynamic dataset. Our proposal relies
on both a linear octree representation with Morton codes and an
efficient integral computation on GPU.  The source code and additional
material (video\ldots) are available on the project website
(\url{https://github.com/dcoeurjo/ICVT}).

>>>>

\begin{figure}[t]{\small
    \begin{center}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/integral2D_2}
          \put(42,28){$\times$}
          \put(46,31){$\vx$}
          \put(55,45){$B_R(\vx)$}
          \put(1,10){$\dS$}
          \put(20,10){$\Shape$}
      \end{overpic}}~~~~~~
      \subfigure[]{\begin{overpic}[width=5cm]{figs/notation_2}
          \put(44.5,25){$h \cdot \vz$}
          \put(55,45){$B_R(\vx)$}
          \put(42,28.5){$\times$}
          \put(45,32.5){$\vx$}
          \put(45,16.5){$h$}
          %\put(88,39){$\Proj(p)$}
          \put(3,11){$\partial \Shape$}
          \put(15,7){$\partial \Body{\DigShape}{h}$}
      \end{overpic}}
      \vspace{-0.4cm}
    \end{center}}
    \caption{Integral invariant computation $(a)$ and
      notations $(b)$ in dimension 2 \cite{CVIU2014}.\label{fig:notations}}
\end{figure}

@@

\begin{figure}[!htbp]
  \begin{center}
    \subfigure{}{\includegraphics[width=\textwidth]{figs/morton}}
    %%\subfigure{}{\includegraphics[height=0.17\textheight]{figs/partitioning}}
    %%\subfigure{}{\includegraphics[height=0.17\textheight]{figs/quadtree_activefront}}
  \end{center}
\caption{Morton codes associated to cells of a linear quatree. Each
   morton code of a child cell is obtained by adding a suffix to its
  parent code \emph{(left)}. The adaptive representation consists of
  quadtree cells whose depth is view point dependent
  \emph{(middle)}. Finally, adaptive Marching-Cubes is used to
  generate the triangulation \emph{(right)}.}
\label{fig_quadtree_partitionning}
\end{figure}

@@

As illustrated in Figure
\ref{fig_quadtree_partitionning}-\emph{middle}, we propose a viewpoint
dependent criterion to decide if a cell needs to be refined: the
closer we are to the camera, the finer the cells are. In addition to
the distance criterion, we define an angular criterion to prevent refining cells outside the camera frustrum. Using a criterion based on the distance between a cell and the camera is well suited to GPU since it can be evaluated independently on all cells. Hence, we can efficiently determine if a cell must be subdivided or not. Figure \ref{fig_lod_octree}
illustrates the level of details (LoD for short) criterion we
use. More precisely, in dimension 2, if $\alpha$ denotes the viewing
angle, an object at distance $d$ from the camera with diameter greater than
 $2\cdot d\cdot\tan(\alpha)$
 is projected into more than one pixel
(see Figure~\ref{fig_lod_octree}-\emph{left}). Hence, the distance
criterion is based on the ratio (\emph{visibility ratio} in the
following)between the cell diameter $l(c)$ (power of 2 depending on
the depth), and
$2\cdot d_c\cdot\tan(\alpha)$, $d_c$ being the distance between the cell center and the camera. For a given cell $c$,
split and merge decision are based on this visibility ratio:
\begin{itemize}
\item $c$ is split if its children cells have a visibility ratio
  greater than  constant $k$\,;
\item $c$ and its sibling cells are merged if their parent cell $c'$
  have av visibility ratio lower than $k$\;
\item otherwise, the cell $c$ stays for the next frame.
\end{itemize}
%
\begin{figure}[!h]
  \begin{center}
    \begin{overpic}[width=4cm]{figs/criterion}
      \put(33,40){$\alpha$}
      \put(50,45){$d_c$}
      \put(95,65){$l(c)$}
    \end{overpic}~~~~~
    \includegraphics[width=4cm]{viewlod2_small}~~~~~
    \includegraphics[width=3.5cm]{subdivision}
  \end{center}
  \caption{Notations \emph{(left)} and adaptive meshing in dimension 2 using the LoD distance and
    angular criterion \emph{(middle)}. Based on LoD criterion
    evaluation, the cell buffer is
  updated between each frame \emph{(right)}.}
  \label{fig_lod_octree}
\end{figure}
%
In addition to the distance criterion, we add an angular term in order
to prevent subdivision of cells far from the camera frustrum (see
Figure~\ref{fig_lod_octree}-\emph{middle}). We do not go into the
details but please note that stay, split and merge decisions of all cells are computed in
parallel from all cell morton codes.

@@

To implement the integration on $B_R(\vx)\cap\Shape$\todo{Check ball
  radii w.r.t. notations in sect2}
(Fig.~\ref{fig:approx}-$a$), a naive approach consists in scanning all
digital points inside such domain (Fig.~\ref{fig:approx}-$b$) and then
to estimate the geometrical moment as the sum of the geometrical
moments of each elementary cubes lying in the intersection. On GPU,
\emph{mipmap} textures are used to obtain a multi-resolution information: if the
input binary object is stored in a 3D texture, GPU hardware can
construct a multi-resolution pyramid such that 8 neighboring voxel
intensities at level $l$ are averaged to define the voxel value at
level $l+1$. If the level 0 corresponds to the binary input object, at
a given level $l$, a voxel $(x,y,z)$ contains the fraction of $\DSh$
belonging to the cube of center $(x,y,z)$ and edge length $2^l$. As a
consequence, we can approximate the volume of $B_R(x)\cap\Shape$ by
considering mipmap values at a given resolution $l$
(Fig.~\ref{fig:approx}-$c$). In this case, errors only occurs for
cells lying inside $\Shape$ (with density 1) not entirely covered by
$B_R(\vx)$.
\begin{figure}
  \begin{center}
    \subfigure[]{\includegraphics[width=4cm]{figs/approx1}}
    \subfigure[]{\includegraphics[width=4cm]{figs/approx-reg-1}}
    \subfigure[]{\includegraphics[width=4cm]{figs/approx-reg-2}}
  \end{center}
  \caption{Integral computation with an area estimation from pixel
    counting $(b)$ or from densities (represented by different
    grayvalues) in a mipmap approximation of $\Shape$.}
  \label{fig:approx}
\end{figure}

Using the hierachical nature of the mipmap texture, we could also
consider hierarchical decompositions of either $B_R(\vx)$
(Fig.~\ref{fig:approx2}-$a$) or $B_R(\vx)\cap\Shape$
(Fig.~\ref{fig:approx2}-$b$). The idea would be to decompose the
regions into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities. Note that the scenario in
Fig.~\ref{fig:approx2}-$b$, is clearly the best approach: the
volume is exact  (\wrt the regular grid case) since geometrical moments
are additive quantities, and the number of texture fetches is  optimal
at this point.

\begin{figure}
  \begin{center}
    \subfigure[]{\includegraphics[width=5cm]{figs/approx-bh}}
    \subfigure[]{\includegraphics[width=5cm]{figs/approx-ada}}
  \end{center}
  \caption{Adaptive approximation of the integration using a
    hierarchical decomposition of the ball $B_R(\vx)$ $(a)$, or a decomposition of $B_R(\vx)\cap\Shape$.}
  \label{fig:approx2}
\end{figure}

@@

\paragraph{Hierarchical Decomposition.}  As discussed above, best
results in terms of texture fetch can be obtained by the approach
presented in Fig.~\ref{fig:approx2}-$b$. At each vertex, the
computation algorithm would consist in a recursive decomposition of
the domain such that a cell is subdivided if it does not belong to
$B_R(\vx)\cap\Shape$. However, on GPU hardware, this approach is
highly inefficient since the hardware optimizes the parallelism only
when the micro-program described in the shader has a predictive
execution flow. Hence, implementing the recursive algorithm (or its
de-recursified version) requires a lot of branches (conditional
\emph{if} tests) in the flow. Instead of implementing this approach,
we have considered a fast hierarchical approximation following
Fig.~\ref{fig:approx2}-$a$: for a given radius $R$, we precompute a
hierachical decomposition of $B_R$. Such decomposition is made of
octree cells at different resolutions. At a given MC vertex $\vx$, the
algorithm becomes simple since we just scan  the ball cells and estimate
the area (or other moments) from the mipmap values associated to this
cell. Note that cells in the $B_R$ decomposition may not be aligned
wuth mipmap cells. However, we can use GPU mipmap functionalities
which interpolates mipmap values if we probe at non-discrete
positions. For a given radius $R$, the expected number of cells on
$B_r$ is in $O(\log{R})$.

@@

\begin{table}
  \begin{center}
    \begin{tabular}{@{}lrrrrrr@{}}
      \toprule
    Quantites  & H & $l_4$ & $l_3$ & $l_2$ & $l_1$ & $l_0$\\
      \midrule
      Number of texture fetches & & & & & & \\
      Time (in ms) & & & & & & \\
      $L_\infty$ error (\wrt $l_0$) & & & & & & \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{object, size, radius, full res, num of vertices (timings
    only consider the curvature estimation step).\label{tab:full-res-stat}}
\end{table}

@@

@@

In this article, we have propose a full data parallel framework on GPU
hardware which combines an adaptive isosurface construction from
digital data with a curvature tensor estimation at each vertex. Using
this approach, we can explore in realtime different curvature
measurements (mean, Gaussian, principal directions) with different
ball radii on potentially large dynamic dataset. Our proposal relies
on both a linear octree representation with Morton codes and an
efficient integral computation on GPU.  The source code and additional
material (video\ldots) are available on the project website
(\url{https://github.com/dcoeurjo/ICTV}).
